---
title: "Data Science in R: Assignment5_Project2 "
author: "Steven Tran, Matthew David, Praveen Kenderla"
date: "March 19, 2018"
output:
    ioslides_presentation: null
    smaller: yes
    slidy_presentation: null
    df_print: paged
subtitle: R project 1
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
```

## Obtaining Data{.smaller}

```{r}
library(tidyverse)
library(stringr)
library(dplyr)
library(ggplot2)
url1 <- "http://www.ndbc.noaa.gov/view_text_file.php?filename=mlrf1h"
url2 <- ".txt.gz&dir=data/historical/stdmet/"
years <- c(1987:2016)
urls <- str_c(url1, years, url2, sep = "")
filenames <- str_c("mr", years, sep = "")
N <- length(urls)
```


## Formatting Data{.smaller} 
```{r}
for (i in 1:N){
  suppressMessages(
    assign(filenames[i], read_table(urls[i], col_names = TRUE))
  )
  file <- get(filenames[i])
  colnames(file)[1] <-"YYYY"
  if(!is.numeric(file[1,1])){
    file <- file[2:nrow(file),]
    file$YYYY <- as.numeric(file$YYYY)
  }
  file$YYYY <- ifelse(as.numeric(file$YYYY%/%100)<1, file$YYYY+1900, file$YYYY) #we add 1900 to the 2 digit years to match YYYY format
  file <- file %>% filter(hh==12)

## Combining Data{.smaller}
 if(i == 1){
    MR <- file %>% select(YYYY, MM, DD, hh, ATMP, WTMP)
}  
else{
    MR <- rbind.data.frame(MR, file %>% select(YYYY, MM, DD, hh, ATMP, WTMP))
  }
}
```

## Making Data Numeric{.smaller}

```{r}
MR$MM <- as.numeric(MR$MM)
MR$DD <- as.numeric(MR$DD)
MR$hh <- as.numeric(MR$hh)
MR$ATMP <- as.numeric(MR$ATMP)
MR$WTMP <- as.numeric(MR$WTMP)
MR <- MR %>% 
  filter(ATMP<99) %>% 
  filter(WTMP<99)
```

## Making Time Series for Air Temperature{.smaller}

``` {r fig.height = 2}
A <- MR %>% 
  select(YYYY,ATMP, MM, DD) %>% 
  mutate(date=as.Date(paste(as.character(MM), "/", as.character(DD), "/", as.character(YYYY), sep=""), format = "%m/%d/%Y"))
ggplot(data=A, aes(x=date, y=ATMP))+
  geom_line()+
  xlab("Year")+
  ylab("Air Temperature")+
  ggtitle("Air Temperature Over Time")+
  scale_x_date(date_labels = "%Y")  
```

## Making Time Series for Water Temperature{.smaller}

``` {r fig.height = 2}
W <- MR %>% 
  select(YYYY,WTMP, MM, DD) %>% 
  mutate(date=as.Date(paste(as.character(MM), "/", as.character(DD), "/", as.character(YYYY), sep=""), format = "%m/%d/%Y"))
ggplot(data=W, aes(x=date, y=WTMP))+
  geom_line()+
  xlab("Year")+
  ylab("Water Temperature")+
  ggtitle("Water Temperature Over Time")+
  scale_x_date(date_labels = "%Y")
```

## correlation between Air and Water Temperatures{.smaller}

``` {r}
x <- cor(A %>% select(ATMP), W %>% select(WTMP), method="pearson")
print(paste("testing the pearson correlation between the Air and Water temperature 
            time series for the 12:00 PM reveals a correlation of ",x, ". This means
            that the changes in temperatures of air and water over the years at the
            location of the buoy somewhat strongly correlate."))

```

## Testing Mean Air Temperature{.smaller}

```{r}
MR1987.ATMP <- as.vector((MR %>% filter(YYYY==1987))[["ATMP"]])
MR2016.ATMP <- as.vector((MR %>% filter(YYYY==2016))[["ATMP"]])
t.test(MR1987.ATMP, MR2016.ATMP)  
```

## Testing Mean Water Temperature{.smaller}

``` {r}
MR1987.WTMP <- as.vector((MR %>% filter(YYYY==1987))[["WTMP"]])
MR2016.WTMP <- as.vector((MR %>% filter(YYYY==2016))[["WTMP"]])
t.test(MR1987.WTMP, MR2016.WTMP)
```


## Part 2 {.flexbox .vcenter}

## Obtaining the data{.smaller}
``` {r}
library(readxl)
veg.1 <- read_xlsx("veg1.xlsx")
```

## Changing the Column Names{.smaller}

```{r}

cnames.1 <- colnames(veg.1)
c <- apply(veg.1, 2, n_distinct)
e <- names(c[c>1])
veg.2 <- select(veg.1, e)
veg.3 <- dplyr::rename(veg.2, 
                       Geo = `Geo Level`, 
                       State = `State ANSI`,
                       Data = `Data Item`,
                       Category = `Domain Category`)

```

## Cleaning the data{.smaller}
``` {r}
yy <- separate(veg.3, Category, into = c("label", "quant"), sep=",")
ru <- filter(yy, label=="RESTRICTED USE CHEMICAL")
ru1 <- ru %>% select(label, quant, Value) %>% unique()
CAS <- t(as.data.frame(strsplit(ru1$quant, "=")))
CAS[,2] <- substr(CAS[,2], 1, nchar(CAS[,2])-1)
ru2 <- cbind(ru1, CAS[,2])
row.names(ru2) <- 1:284
colnames(ru2) <- c(colnames(ru2)[1:length(colnames(ru2))-1], "CAS")
```


## {.smaller}

``` {r}
ru3 <- ru1 %>% separate (quant, c("ChemType", "ChemName"), sep = ":") %>%
  mutate (Value = str_extract(Value, "^\\w+"))

ru4 <- unique(ru3[, c("label", "ChemType" , "ChemName", "Value")]) %>%
  na.omit(ru8) # Only takes the unique Chemical Names and corresponding values;

ru5 <- ru4[1:5, ]  # Clean the Value column NA's and 
                    # makes a table with corresponding chemical names
                    # for first 5 rows
```

## Toxicity data taken from Website{.smaller}

```{r}
ToixicityData <- tibble (TestedSpecies = c("NorwayRat", "HoneyBee", "FruitMoth", 
                                           "NorwayRat", "BoolWeevil" ),
                   Effectmeasurement = c("Activity_General", "Mortality", "Mortality",
                                         "Mortality", "Mortality"),
                    LD50dose = c(4, 7, 7, 6, 4),
                   Dosage = c(23.3, 0.034,  0.011, 451, 0.0034),
                   Units = c("mg/kg", "ug/ul", "ug/org", "mg/kg", "ug/org"),
                   ObservedDurationiDays = c(0.0625, 1, 1, 1, 3))
# Rows extracted from search: 953, 190, 508, 52, 235
```
-Toxicity data is extrated from the Beta version of ECOTOX website
- only first five chemicals in the ru5 variable are extracted
-Six columns include:
1) On what species test is conducted, 2) what is measured; 
3) Total number of exposure overal, 4) Dosage required to show LD50
5) Units of the dosage used in experiment and 6) Observed duration days

## Final table{.smaller}
- Data shows that very small amount of dosage of these chemicals can show lethal effects
- Most of the cases, lethan effects include Mortality rates. 
- By this data, we can indicate that using these chemicals have severe effect on animal populations. 
 Also, large quantities of these chemicals can effect health of mammals in addition to other animals.

```{r}
ru6 <- as.tibble(c(ru5, ToixicityData))
Part2 <- ru6
Part2
```